---
title: "Thesis"
author: "ENOCK OSEI"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  fig.width = 14,fig.height = 8, size = 16,	message = FALSE, warning = FALSE, comment = NA)
library(tidyverse)
library(gtsummary)
library(summarytools)
library(gt)
library(ggstatsplot)
library(finalfit)
library(flextable)
library(ggpubr)
library(nnet)
library(haven)
library(easystats)
library(readxl)
library(huxtable)
before <- read.csv("bc.csv")
during <- read.csv("dc.csv")
#during <- read.csv("C:\\R Projects\\thesis\\during.csv")
codebook <- read.csv("codebk.csv")
Hmisc::label(before, self = F) <- codebook$label
Hmisc::label(during, self = F) <- codebook$label
#str(before)
```

## Demography
```{r}
before %>% select(age, gender, edu, employ_status, marital_status, household_car_own, motor_bicycle_own, income) %>% 
  tbl_summary() %>% 
  bold_labels() %>% 
  as_flex_table()
```

```{r eval=FALSE, include=FALSE}
thesis %>% select(primary_trips, work_trip, study_trip, shopping_trips, recreational_activity, social_activity, e_learning, teleshopping, primary_trip, distance_traveled, mode_of_transport, reason_of_mode, primary_trip, w_trips, st_trips, sh_trips, rec_ac_trip, soci_ac_trip, elearning, tele_working, tele_shopping, dis_travel, change_mode, transport_mode, fac_inf_mode, fac_res_movement) %>% filter(change_mode!="") %>% tbl_summary() %>% bold_labels() %>% as_flex_table()
```

## Primary Trips Before COvid By Gender
```{r}
sex<-before %>% select(gender, primary_trips) %>% tbl_summary(by = "gender") %>% bold_labels()
ed<-before %>% select(edu, primary_trips) %>% tbl_summary(by = "edu") %>% bold_labels()
tbl_merge(tbls = list(sex, ed), tab_spanner = c("Gender", "Education")) %>% as_flex_table()
```

## Frequency of Primary Trips Before and During Covid
```{r}
bf<-before %>% select(work_trip, study_trip, shopping_trips, recreational_activity, social_activity) %>% tbl_summary()
du<-during %>% select(work_trip, study_trip, shopping_trips, recreational_activity, social_activity) %>% tbl_summary()
tbl_merge(tbls = list(bf, du), tab_spanner = c("Before Covid", "During Covid")) %>% bold_labels() %>% as_flex_table()
```

## Frequency of Online Activities Before and During Covid
```{r}
bf<-before %>% select(e_learning, teleshopping, teleworking) %>% tbl_summary()
du<-during %>% select(e_learning, teleshopping, teleworking) %>% tbl_summary()
tbl_merge(tbls = list(bf, du), tab_spanner = c("Before Covid", "During Covid")) %>% bold_labels() %>% as_flex_table()
```

## Frequency of Primary tripps Before and During Covid by Gender
```{r}
bf<-before %>% select(primary_trips) %>% tbl_summary()
du<-during %>% select(primary_trips) %>% tbl_summary()
tbl_merge(tbls = list(bf, du), tab_spanner = c("Before Covid", "During Covid")) %>% bold_labels() %>% as_flex_table()
```

## Frequency of distance traveled Before and During Covid by Gender
```{r}
bf<-before %>% select(distance_traveled, gender) %>% tbl_summary(by = gender)
du<-during %>% select(distance_traveled, gender) %>% tbl_summary(by = gender)
tbl_merge(tbls = list(bf, du), tab_spanner = c("Before Covid", "During Covid")) %>% bold_labels() %>% as_flex_table()
```

## Frequency of distance traveled Before and During Covid by Household car ownership
```{r}
bf<-before %>% select(household_car_own, distance_traveled) %>% filter(household_car_own!="no")%>% tbl_summary(by = household_car_own)
du<-during %>% select(household_car_own, distance_traveled) %>%filter(household_car_own!="no") %>% tbl_summary(by = household_car_own)
tbl_merge(tbls = list(bf, du), tab_spanner = c("Before Covid", "During Covid")) %>% bold_labels() %>% as_flex_table()
```

```{r eval=FALSE, include=FALSE}
thesis <- read_excel("questionnaire.xlsx")
thesis %>%
  pivot_longer(27:32, names_to = "Risk", values_to = "Response") %>% 
  mutate(
    Risk = fct_recode(Risk, 
                     "Risk of Covid in Private car" = "risk_private_car",
                  "Risk of Covid in Trotro" = "risk_trotro",
                  "Risk of Covid in Uber or Bolt" = "risk_uber_bolt",
                   "Risk of Covid on Bicycle" = "risk_bicycle",
                   "Risk of Covid in Moto Bicycle" = "risk_motorbike",
                   "Risk of Covid in Walking" = "risk_walking")) %>% 
      select(Risk, Response) %>% filter(Response !="") %>% 
  tbl_cross(row = Risk, col = Response, margin = "column", percent = "cell") %>% 
  modify_caption("**Table 1**") %>% 
  as_flex_table()
```

## Perceived Risk of infection (Graph)
```{r}
before <- read.csv("bc.csv")
before %>%
  pivot_longer(17:22, names_to = "Risk", values_to = "Response") %>% 
  mutate(
    Risk = fct_recode(Risk, 
                     "Risk of Covid in Private car" = "private_car",
                  "Risk of Covid in Trotro" = "trotro",
                  "Risk of Covid in Uber or Bolt" = "uber_bolt",
                   "Risk of Covid on Bicycle" = "bicycle",
                   "Risk of Covid in Moto Bicycle" = "motorbike",
                   "Risk of Covid in Walking" = "walking")) %>% 
  filter(Response !="6") %>% 
   ggbarstats(x = Response, y = Risk, ylab = "status", title = "Respondent's Response to Risks of Traveling", bf.message = F, proportion.test = F,   results.subtitle = FALSE) + 
  #scale_fill_lancet() +
  theme(
    legend.position = "bottom"
  ) +
  coord_flip()
```

## Multinomial Logistic Regression
# @ before covid


## We now split the dataset into train and test using "sample_frac()" function from {dplyr} package
# Using sample_frac to create 70 - 30 slipt into test and train
```{r eval=FALSE, include=FALSE}
train_before <- sample_frac(before, 0.8)
sample_id <- as.numeric(rownames(train)) # rownames() returns character so as.numeric
test_before <- before[-sample_id,]
```

# Another way of spliting the data
```{r}
#Splitting the data using a function from dplyr package
library(caret)

index <- createDataPartition(before$mode_transport, p = .80, list = FALSE)
train <- before[index,]
test <- before[-index,]
```

```{r}
before$mode_transport <- as.factor(before$mode_transport)
```

# Benchmark or reference
```{r}
before$mode_transport <- relevel(before$mode_transport, ref = "Non_motorized_transport")
```

## model 
```{r}
modl <- multinom(mode_transport ~ primary_trips + gender + age + edu_level + marital_status + employ_status, data = before)
summary(modl)
```

# Odds
```{r}
exp(coef(modl))
```

## p-value
```{r}
zvalues <- summary(modl)$coefficients/summary(modl)$standard.error
p <- (1-pnorm(abs(zvalues), 0, 1))*2
p
```

# Predicting and Accuracy the class on train dataset.
```{r}
# Predicting the values for train dataset
train$predicted <- predict(modl, newdata = train, "class")
 
# Building classification table
tab <- table(train$mode_transport, train$predicted)
 
# Calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(tab))/sum(tab)) * 100, 2)
```

## Predicting and Accuracy the class on test dataset.
```{r}
# Predicting the class for test dataset
test$Predicted <- predict(modl, newdata = test, "class")

# Building classification table
tab <- table(test$mode_transport, test$Predicted)

# Calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(tab))/sum(tab)) * 100, 2)
```

# Prediction
```{r}
head(predict(modl, before, type = "prob"))## prediction for the first 6 obs. Remove head to see all
```

# Misclassification error
```{r}
mce <- table(predict(modl), before$mode_transport)
mce
```
263 respondents were classified as Paratransiters and the model confirms that is true. However, 2 of them were classified as Ride_hailing but the model says they are Private_transporters. etc                

# Misclassification Error Percentage
```{r}
round((1 - sum(diag(mce))/sum(mce)) * 100 ,2)
```

# @ during covid

# Another way of spliting the data
```{r}
#Splitting the data using a function from dplyr package
library(caret)

index <- createDataPartition(during$transport_mode, p = .80, list = FALSE)
train <- during[index,]
test <- during[-index,]
```

```{r}
during$transport_mode <- as.factor(during$transport_mode)
```

# Benchmark or reference
```{r}
during$transport_mode <- relevel(during$transport_mode, ref = "Non motorized transport")
```

## model 
```{r}
modll <- multinom(transport_mode ~ primary_trips + gender + age + edu_level + marital_status + employ_status, data = during)
summary(modll)
```

```{r}
library(finalfit)
library(kableExtra)
explanatory = c("primary_trips", "gender", "age", "edu_level", "marital_status", "employ_status")
dependent = "transport_mode"
during %>%
 finalfit(dependent, explanatory,
          metrics = TRUE, keep_models = TRUE) -> tb
knitr::kable(tb, row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
```

#Odds
```{r}
exp(coef(modll))
```

## p-value
```{r}
zvalues <- summary(modll)$coefficients/summary(modll)$standard.error
p <- (1-pnorm(abs(zvalues), 0, 1))*2
p
```
# Accuracy
```{r}
# Predicting the values for train dataset
during$precticed <- predict(modll, newdata = during, "class")
 
# Building classification table
ctable <- table(during$transport_mode, during$precticed)
 
# Calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(ctable))/sum(ctable))*100,2)
```

## Predicting and Accuracy the class on test dataset.
```{r}
# Predicting the class for test dataset
test$Predicted <- predict(modll, newdata = test, "class")

# Building classification table
tab <- table(test$transport_mode, test$Predicted)

# Calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(tab))/sum(tab)) * 100, 2)
```
We were able to achieve 63.29% accuracy in the test dataset and this number is very close to train, and thus we conclude that the model is good and is also stable.

# Prediction
```{r}
head(predict(modll, during, type = "prob")) ## prediction for the first 6 obs. Remove head to see all
```

# Predicted Probabilities 
```{r eval=FALSE, include=FALSE}
head(pp <- fitted(modll))
```

# Misclassification error
```{r}
mce <- table(predict(modll), during$transport_mode)
mce
```
181 respondents were classified as Paratransiters and the model confirms that is true. However, 2 of them were classified as Paratransiters but the model says they are Non motorized transporters. 

# Misclassification Error Percentage
```{r}
round((1 - sum(diag(mce))/sum(mce))*100,2)
```
